Partial RDA

### 2019-05-10 partialrda.Rmd
Library
Genus + core

```{r libraries, cache = FALSE}
library(dplyr)
library(vegan)
library(tibble)
```

```{r data, eval = FALSE}
rda.abundances  <- as.matrix(t(microbiome::abundances(microbiome::transform(pseq.species, "clr"))))
permutest(rda.species, method="direct", perm=100)
rda.species <- rda(rda.abundances ~ MAP + Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
                   data = meta(pseq.species))
saveRDS(rda.species, file = "rda.specids.rds")
```

```{r load}
rda.species <- readRDS("rda.specids.rds")
```

```{r variance}
data.frame(Inertia = c(
               Conditioned = summary(rda.species)$partial.chi,
               Constrained = summary(rda.species)$constr.chi,
               Unconstrained = summary(rda.species)$unconst.chi)) %>%
    tibble::rownames_to_column("Variance") %>%
    mutate(Proportion = Inertia/sum(Inertia)) %>%
    mutate_at(2:2, funs(round(., 1))) %>%
    mutate_at(3:3, funs(round(., 5))) %>%
    kable
```

```{r direct}
permutest(rda.species, method="direct", perm=100)
```

```{r axis}
permutest(rda.species, method="axis", perm=100)
```

```{r term}
permutest(rda.species, method="term", perm=100)
```

```{r rda plot}
jpeg(file="cache/rda.jpg", res = 300, width = 20, height = 20, units = 'cm')
plot(rda.species)
dev.off()
```

```{r loading, eval = FALSE}
loadings <- as.data.frame(rda.species$CA$v.eig[,1:2])

rda.species %>% str

loadings %>% str
```

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(vegan)
library(broom)
library(tibble)
library(knitr)
library(phyloseq)
library(microbiome)
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

</details>

```{r functions}
myresults.table <- function(model) {
    data.frame(Inertia = c(
                   Conditioned = summary(model)$partial.chi,
                   Constrained = summary(model)$constr.chi,
                   Unconstrained = summary(model)$unconst.chi)) %>%
        tibble::rownames_to_column("Variance") %>%
        mutate(Proportion = Inertia/sum(Inertia)) %>%
        mutate_at(2:2, funs(round(., 1))) %>%
        mutate_at(3:3, funs(round(., 5))) %>%
        kable
}

anotab <- function(x) {
    EPS <- sqrt(.Machine$double.eps)
    Pval <- (colSums(sweep(x$F.perm, 2, x$F.0 - EPS, ">=")) + 1)/(x$nperm + 1)
    anotab <- data.frame(x$df, x$chi, c(x$F.0, NA), c(Pval, NA))
    colnames(anotab) <- c("Df", "Inertia", "F", "Pr(>F)")
    rownames(anotab) <- c(x$termlabels, "Residual")
    class(anotab) <- c("anova", "data.frame")
    anotab 
}
```

Loading phyloseq object

```{r data}
pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
```

Updating phyloseq objects

```{r updating phyloseq objects}
phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(filter.phenotype.data(pseq.genus))
phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
```

```{r data}
rda.abundances.species  <- microbiome::transform(pseq.species, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()

rda.abundances.genus  <- microbiome::transform(pseq.genus, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()

rda.abundances.genus.core  <- microbiome::transform(pseq.genus.core, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()
```

```{r models}
rda.species <- rda(rda.abundances.species ~ MAP +
                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
                   data = meta(pseq.species))
rda.genus <- rda(rda.abundances.genus ~ MAP +
                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
                 data = meta(pseq.genus))
rda.genus.core <- rda(rda.abundances.genus.core ~ MAP +
                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
                   data = meta(pseq.genus.core))

```

```{r myresults species}
myresults.table(rda.species)
```

```{r myresults genus}
myresults.table(rda.genus)
```

```{r myresults genus core}
myresults.table(rda.genus.core)
```

# ANOVAs

Species

```{r anova species}
anova.species <- permutest(rda.species, method="reduced", perm=1)
anova.species %>% anotab %>% kable
```

Genus

```{r anova genus}
anova.genus <- permutest(rda.genus, method="reduced", perm=1)
anova.genus %>% anotab %>% kable
```

```{r anova genus core}
anova.genus.core <- permutest(rda.genus.core, method="reduced", perm=1)
anova.genus.core %>% anotab %>% kable
```

# Plots

Species

```{r rda plot species}
plot(rda.species)
```

Genus

```{r rda plot species}
plot(rda.genus)
```

Genus core

```{r rda plot species}
plot(rda.genus.core)
```
